---
title: "Terminology in Statistics"
author: "Niall McGuinness"
date: "2022-11-30"
output:
  html_document:
    toc: TRUE
    number_sections: FALSE
    toc_float: TRUE
    df_print: paged
    fig_caption: TRUE
    code_folding: hide
    theme: cerulean
  pdf_document:
    toc: TRUE
  word_document:
    toc: TRUE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Useful Links

* [Color in R](https://www.programiz.com/r/colors#:~:text=We%20define%20a%20color%20as,and%20%23000000%20would%20be%20black.)
* [R Markdown Cookbook](https://bookdown.org/yihui/rmarkdown-cookbook/)
* [R Markdown Cheatsheet](https://www.rstudio.com/wp-content/uploads/2015/02/rmarkdown-cheatsheet.pdf)
* [R Markdown Theme Gallery](https://www.datadreaming.org/post/r-markdown-theme-gallery/)
* [An overview of color names in R](https://r-graph-gall
   ery.com/42-colors-names.html)
* [Top 7 Packages for Making Beautiful Tables in R](https://towardsdatascience.com/top-7-packages-for-making-beautiful-tables-in-r-7683d054e541)
* [Beautiful tables in R with gtExtras](https://themockup.blog/posts/2022-06-13-gtextras-cran/)
* [Understanding YAML headers](https://zsmith27.github.io/rmarkdown_crash-course/lesson-4-yaml-headers.html)
* [Setting up your R markdown YAML Header](https://zsmith27.github.io/rmarkdown_crash-course/lesson-4-yaml-headers.html)

# What is markdown?

## How can we add headings?

## How can we add subheadings?

This is in **bold**. This is some *words* in italic.

This is a bulleted list:
* Player - data
* Customer - data

This is a numbered list:
1. Something
2. Something else

Can we add a link to the [Dkit]("www.dkit.ie")?

How should we add a table?

Table Header | Second Header
------------- | -------------
Table Cell | Cell 2
Cell 3 | Cell 4 

# What are the key terms in statistics?

## What types of variables exist in a game-related survey?

* Name - character - "John"
* Gender - character - m/f/o/u
* Age - integer - 45, 32
* Satisfied - integer - 1->5, unsatisfied/satisfied
* Time - float-point - 2.25hrs
* Price - float-point - 2.25hrs

### Define categories for the examples above

* Numerical - continuous, discrete e.g. height, age
* Categorical - nominal (name) e.g. "brown", "male", "satisfied"

## What is a statistic? 

* If we derive a value/variable from a sample we call it a statistic
e.g. average age of a 100 people then we have a **sample statistic**.

* If we derive a value/variable from the population we call it a parameter
e.g. census gives us average age of the population then we have a **population parameter**.

statistics -> samples
parameters -> population

### What is an example of a statistic?

```{r echo=FALSE}
# answer in range 1-5
satisfaction_data <- c(3.2,2,5.1,1,4,5.0,3,2.2,1,4,5.1,1,2.8,3.3,1)
satisfaction_data

# mean or average
avg_satisfaction <- mean(satisfaction_data)
avg_satisfaction

# min or max
min_satisfaction <- min(satisfaction_data)
min_satisfaction

# mode
# mode_satisfaction <- mode(satisfaction_data)
# mode_satisfaction

# median (we sort the numbers and we get the centre number)
simpler_data <- c(1,2,3,4,5,6)

median_satisfaction <- median(simpler_data)
median_satisfaction
```

This average of the data is `r round(avg_satisfaction, 2)` satisfaction points.


# What can we plot with R Studio?

Well, we can generate plots really quickly from CSV file, or from remote RBDMS data, or from CSV file on web page.

### Histogram

```{r echo=FALSE}
hist(satisfaction_data,
     main = "This is the title",
     xlab = "Satisfaction",
     ylab = "# Satisfied",
     col = rgb(1, 1, 0))
```

### Barplot

```{r}
gamer_survey_data <- read.csv("amalgamated_game_survey_250_2022.csv")
barplot(sort(table(gamer_survey_data$gender), decreasing = FALSE),
        main = "My title",
        xlab = "Breakdown of Age by Gender",
        ylab = "User Number",
        ylim = c(0, 100),
        col=c("red", "coral", "blue", "yellow"))

```


### Boxplot
One box plot...
```{r}
boxplot(gamer_survey_data$age, main = "My boxplot title")
```

Multiple box plots...
```{r}
boxplot(gamer_survey_data$age~gamer_survey_data$gender, main = "My multi-boxplot title")
```

# Can we make tables of our data?

```{r}
# show a "prettier" paged table
rmarkdown::paged_table(gamer_survey_data)

# show a "prettier" paged table but ONLY first 8 rows
head_data <- head(gamer_survey_data, 8)
rmarkdown::paged_table(head_data)

# show a "prettier" paged table using kable() with caption and decimal digit max length and alignment = centre
knitr::kable(gamer_survey_data,
  align = c('r'), # a vector with 'c', 'l', or 'r' values
  digits = 2,
  caption = "A caption for your table here")
```
## Can we make a subset of the data to show in a table?

## Can we filter the data to show in a table?




















